Optimal (Multiway) Spatial Joins

### 2024-05-25 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\out{\mathrm{OUT}}

@@

Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?} A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$. A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}. 
    \label{eqn:cross}
}

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$. Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

@@

We 
conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.

@@

We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-1}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, R_2, ..., R_{k-1}, H', V)$. We prove in Appendix~\ref{app:hv}:

@@

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\out{\mathrm{OUT}}

@@

%Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?}

Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$. A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}. 
    \label{eqn:cross}
}
Let $R$ be a set of rectangles. Given a horizontal segment, we define 
\myeqn{
    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}. 
    \label{eqn:contained}
}

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ (defined in \eqref{eqn:contained}), where the rectangles are sorted by their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

@@

We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.

@@

We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-1}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, ..., R_{k-1}, H', V)$. We prove in Appendix~\ref{app:hv}:

@@

The $(k-1)$-SJ can be settled in $F_{k-1}(n, \out)$ time.

\extraspacing {\bf Generating $\bm{\J^*}$.}
